Partial Least Squares Regression with Multiple Functional and Scalar Predictors

### 2025-06-28 manuscript/main.tex
started abandoning the block diagonal notation

	\subsubsection{Finite-basis approximation of the predictors and regression coefficients}\label{sec: finite basis approximation}
	\paragraph{Basis functions.}
	For $k = 1, \ldots, K$, let us denote the system of $M$ basis functions used to approximate the $k$-th functional predictor as 
	\begin{equation*}
		\mathbf{b}^{(k)}(t_k)=(b^{(k)}_{1}(t_k), \ldots, b^{(k)}_{M}(t_k))^\top,~\text{where}~t_k \in \mathcal{T}_k.   
	\end{equation*}
	In our proposed hybrid PLS regression algorithm, this basis system is used to approximate both the predictors and the PLS components. It also serves as a regularization tool to borrow strength and prevent overfitting.

	\paragraph{Basis approximation of the hybrid predictor.}
	Then we can approximate the functional predictor observations as
	\begin{equation*}
		X_i^{(k)}(t_k) = \sum_{m=1}^M c_{im}^{(k)} b^{(k)}_{m}(t_k)=\mathbf{c}_i^{(k)T}\mathbf{b}^{(k)}(t_k), \;\; t_k \in \mathcal{T}_k, \;\; k=1, \ldots, K, \;\; i=1, \ldots, n,
	\end{equation*}
	where $\mathbf{c}_i^{(k)} = (c_{i1}^{(k)}, \ldots, c_{iM}^{(k)})^\top$ is a $M$-dimensional vector of basis coefficients corresponding to the $k$th functional predictor. This basis approximation can be expressed collectively across $n$ observations as: $\widetilde{\mathbf{X}^{(k)}}(t_k) = \widetilde{C}^{(k)} \mathbf{b}^{(k)}(t_k)$, where $\widetilde{\mathbf{X}}^{(k)}(t_k)=(X_1^{(k)}(t_k), \ldots, X_n^{(k)}(t_k))^\top \in \mathbb{R}^n$, and $\widetilde{C}^{(k)} = (\mathbf{c}_1^{(k)} \cdots \mathbf{c}_n^{(k)})^\top \in \mathbb{R}^{n \times M}$. This notation can be further extended to simultaneously express the basis expansions of $K$ functional predictors as 
	\begin{equation}
		\widetilde{X}(\mathbf{t}) = \widetilde{C}B(\mathbf{t}), \quad t \in \mathcal{T},
		\label{eq: basis approximation of X}
	\end{equation}
	where $\widetilde{X}(\mathbf{t}) = [\widetilde{\mathbf{X}}^{(1)}(t_1), \cdots, \widetilde{\mathbf{X}}^{(K)}(t_K)] \in \mathbb{R}^{n \times K}$, $\widetilde{C} = [\widetilde{C}^{(1)}, \cdots, \widetilde{C}^{(K)}] \in \mathbb{R}^{n \times MK}$, and 
	$B(\mathbf{t}) = \mathrm{blkdiag}[\mathbf{b}^{(1)}(t_1), \cdots, \mathbf{b}^{(K)}(t_K)] \in \mathbb{R}^{MK \times K}$.
	The number of basis functions $M$ for each functional predictor can be set to a sufficiently large value (e.g., 15 or 20) to incorporate a rich variety of patterns that functional predictors may exhibit. Note that there is no need to carefully choose different $M$ values for different functional predictors, as an appropriate penalization scheme will be introduced in the later section to regularize each functional predictor to achieve a desired smoothness (see Section \ref{sec: Regularization of PLS Components}).

	Furthermore, let $\widetilde{Z} = [\mathbf{Z}_1, \cdots, \mathbf{Z}_n] \in \mathbb{R}^{n \times p}$ denote $n \times p$ matrix of the scalar predictors (observed version when $l=1$, or residualized version when $l \ge 2$), and let $\widetilde{W}[\mathbf{t}]=[\widetilde{X}(\mathbf{t}), \widetilde{Z}]$ denote $n \times (K+p)$ hybrid predictor matrix that stacks $\widetilde{X}(\mathbf{t})$ and $\widetilde{Z}$ as columns. Then we can also write the hybrid predictor, whose functional part is approximated using the basis functions, as a form of linear transformation similar to \eqref{eq: basis approximation of X}:
	\begin{equation}\label{eq: basis approximation of W}
		\widetilde{W}[\mathbf{t}] 
		=  
		\widetilde{ \mathbf{C} }^\ast B^*[\mathbf{t}], \quad t \in \mathcal{T}.
	\end{equation}
	where  $B^*[\mathbf{t}] = \mathrm{blkdiag}(B(\mathbf{t}), I_p) \in \mathbb{R}^{(MK+p) \times (K+p)}$ and $\widetilde{ \mathbf{C} }^\ast=[\widetilde{C}, \widetilde{Z}] \in \mathbb{R}^{n \times (MK+p)}$.

	\paragraph{Basis approximation of the PLS component.}
	We can also approximate the functional part of the PLS component in terms of the same basis functions: $\psi^{(k)}(t_k) = \sum_{m=1}^M a^{(k)}_{m} b^{(k)}_m(t_k) = \mathbf{a}^{(k)\top} \mathbf{b}^{(k)}(t_k)$ and $\psi(\mathbf{t}) = B(\mathbf{t})^\top \mathbf{a} \in \mathbb{R}^K$, where $\mathbf{a}^{(k)} = (a^{(k)}_{1}, \ldots, a^{(k)}_{M})^\top \in \mathbb{R}^M$, and $\mathbf{a}=(\mathbf{a}^{(1)\top}, \ldots, \mathbf{a}^{(K)\top})^\top \in \mathbb{R}^{MK}$, so that $\boldsymbol{\xi}[\mathbf{t}] = (\psi(\mathbf{t})^\top, \boldsymbol{\theta}^\top)^\top = (\mathbf{a}^\top B(\mathbf{t}), \boldsymbol{\theta}^\top)^\top \in \mathbb{R}^{K+p}$. 
	%
	For each $k = 1, \ldots, K$, let $J^{(k)} \in \mathbb{R}^{M \times M}$ denote the Gram matrix of the basis functions associated with the $k$-th functional predictor, defined by
	\begin{equation}\label{def:gram_basis}
		J^{(k)} = \left[ \int_{\mathcal{T}k} b_m^{(k)}(t) \, b_n^{(k)}(t) , dt \right]_{m,n=1}^M.
	\end{equation}
	Using this definition, we construct the following block-diagonal matrices:
	\begin{equation}\label{def:J_and_J_star}
		J = \mathrm{blkdiag}(J^{(1)}, \cdots, J^{(K)}) \in \mathbb{R}^{MK \times MK}~\text{and}~J^\ast=\mathrm{blkdiag}(J, I_p) \in \mathbb{R}^{(MK+p) \times (MK+p)}.
	\end{equation}

@@

We now describe the strategy for computing the PLS direction at the $l$-th iteration. For simplicity, we omit the iteration index and denote the residualized response by $Y$, with its finite-sample observations denoted by $\mathbf{y} = (y_1, \ldots, y_n) \in \mathbb{R}^n$. Let $\widetilde{W} \in \mathcal{H}$ denote the residualized, basis-approximated hybrid predictor, and let $\widetilde{\mathbf{C}}^\ast \in \mathbb{R}^{n \times (MK+p)}$ denote the basis coefficients of its finite-sample observations. Our strategy finds the direction $\boldsymbol{\xi} \in \mathcal{H}$ that maximizes the empirical covariance between the resulting PLS component $\langle \widetilde{W}, \boldsymbol{\xi} \rangle_{\mathcal{H}}$ and the response.

For practical computation, we first construct the following matrix:
	\begin{equation}\label{def:V_star}
	V^\ast = n^{-2}\begin{bmatrix}
		J \widetilde{C}^\top \mathbf{y}\mathbf{y}^\top \widetilde{C} J & J \widetilde{C}^\top \widetilde{\mathbf{y}} \widetilde{\mathbf{y}}^\top \widetilde{Z} \\
		\widetilde{Z}^\top \widetilde{\mathbf{y}} \widetilde{\mathbf{y}}^\top \widetilde{C} J &   \widetilde{Z}^\top \widetilde{\mathbf{y}} \widetilde{\mathbf{y}}^\top \widetilde{Z}
	\end{bmatrix} = n^{-2} J^\ast \widetilde{C}^{*\top} \widetilde{\mathbf{y}} \widetilde{\mathbf{y}}^\top \widetilde{C}^{*} J^\ast
	\in \mathbb{R}^{(MK+p) \times (MK+p)}.
\end{equation}

%	consider the basis function expansions of the functional predictors $X_i^{(k)}(t_k) = \sum_{m=1}^M c_{im}^{(k)} b^{(k)}_{m}(t_k)$, $i=1, \ldots, n$, and those of the functional parts of the PLS component $\psi^{(k)}(t_k) = \sum_{m=1}^M a^{(k)}_{m} b^{(k)}_m(t_k)$, $k=1, \ldots, K$. Let 
	\begin{proposition} \label{Proposition: generalized rayleigh quotient}

		\jongmin{in the (1,1)th block, should $ \mathbf{y}$ be replaced by $\widetilde{\mathbf{y}} $? }

		and  $\boldsymbol{\xi}^*=(\mathbf{a}^{\top}, \boldsymbol{\theta}^{\top})^\top \in \mathbb{R}^{MK + p}$. Then, the $l$-th PLS component, $\widehat{\boldsymbol{\xi}} \equiv \widehat{\boldsymbol{\xi}}_l$ can be obtained by solving
		\begin{equation}\label{def:eigenproblem_smoothed}
			\argmax \limits_{\boldsymbol{\xi}^* \in \mathbb{R}^{MK+p}} \; \frac{\boldsymbol{\xi}^{*\top} V^\ast \boldsymbol{\xi}^*}{\boldsymbol{\xi}^{*\top} J^\ast \boldsymbol{\xi}^*}, \quad \textit{or equivalently}, \quad \argmax \limits_{\boldsymbol{\xi}^* \in \mathbb{R}^{MK+p}} \;  \boldsymbol{\xi}^{*\top} V^\ast \boldsymbol{\xi}^* \quad \textnormal{subject to}  \quad \boldsymbol{\xi}^{*\top} J^\ast \boldsymbol{\xi}^*=1,
		\end{equation}
		with respect to $\widehat{\boldsymbol{\xi}}^*=(\widehat{\mathbf{a}}^{\top}, \widehat{\boldsymbol{\theta}}^{\top})^\top$, and then setting $\widehat{\boldsymbol{\xi}}[\mathbf{t}] = [\widehat{\psi}(\mathbf{t})^\top, \widehat{\boldsymbol{\theta}}^\top]^\top =  B^*[\mathbf{t}]^\top \widehat{\boldsymbol{\xi}}^*$. The $l$-th PLS score $\widehat{\rho}_{i} \equiv \widehat{\rho}_{il}$ can be obtained as $\widehat{\rho}_i = \langle \mathbf{W}_i, \widehat{\boldsymbol{\xi}} \rangle_\mathcal{H}$.
	\end{proposition}
	The proof of Proposition \ref{Proposition: generalized rayleigh quotient} is given in the Appendix \ref{section:proof:eigenproblem_derivation}.
	Here, $\widehat{\textnormal{Cov}}^2(\langle \mathbf{W}, \boldsymbol{\xi} \rangle_{\mathcal{H}}, Y)$ is expressed as $\boldsymbol{\xi}^{*\top} V^\ast \boldsymbol{\xi}^*$ based on the aforementioned basis approximations. Thus, finding $\boldsymbol{\xi}^*$ that maximizes $\boldsymbol{\xi}^{*\top} V^\ast \boldsymbol{\xi}^*$ amounts to finding $\widehat{\boldsymbol{\xi}}$ that maximizes the sample squared covariance $\widehat{\textnormal{Cov}}^2(\langle \mathbf{W}, \boldsymbol{\xi} \rangle_{\mathcal{H}}, Y)$, which is the first step of the PLS algorithm. The constraint $\boldsymbol{\xi}^{*\top} J^\ast \boldsymbol{\xi}^*=1$ scales the obtained PLS component to have a unit norm, that is, $\langle \widehat{\boldsymbol{\xi}}, \widehat{\boldsymbol{\xi}} \rangle_\mathcal{H}=1$.

	\jongmin{
		Another way of computing \eqref{def:eigenproblem_smoothed}, rather than solving an eigenproblem, is as follows. As shown in the proof in Appendix \ref{section:proof:eigenproblem_derivation}, the empirical covariance is simplified as:
		\begin{align*}
			\widehat{\textnormal{Cov}} (\langle \mathbf{W}, \boldsymbol{\xi} \rangle_{\mathcal{H}}, Y ) &= \frac{1}{n} \sum \limits_{i=1}^n \langle \mathbf{W}_i, \boldsymbol{\xi} \rangle_\mathcal{H} Y_i \\
			&= n^{-1} \widetilde{\mathbf{y}}^\top (\widetilde{C}J\mathbf{a} + \widetilde{Z} \boldsymbol{\theta}).
			\\&=
			\frac{1}{n}\tilde{\mathbf{y}}^\top
			\begin{bmatrix} \widetilde{C}J \; \widetilde{Z} \end{bmatrix} \;      \underbrace{
				\begin{bmatrix} \mathbf{a} \\ \mathbf{\theta} \end{bmatrix}
			}_{:= \boldsymbol{\beta}}.
		\end{align*}
		Let us 
		denote $\mathbf{b} =  \widetilde{\mathbf{y}}^\top \begin{bmatrix} \widetilde{C}J \; \widetilde{Z} \end{bmatrix}$. This is the weighted row sum of $\begin{bmatrix} \widetilde{C}J \; \widetilde{Z} \end{bmatrix}$, weighted by $\widetilde{\mathbf{y}}$.
		Then using the symmetricity of $J^\ast$, we solve a linear system $J^\ast \mathbf{d} =
		\mathbf{b}^\top$ to obtain $\mathbf{d}$. Then we have
		\begin{align*}
			\max_{\sqrt{\boldsymbol{\beta} J^\ast  \boldsymbol{\beta}} = 1}
			\widehat{\textnormal{Cov}} (\langle \mathbf{W}, \boldsymbol{\xi} \rangle_{\mathcal{H}}, Y ) 
			%%%%%%%%%%%%%%%%%%%%%
			&= 
			%%%%%%%%%%%%%%%%%%%%%%%%
			\max_{\sqrt{\boldsymbol{\beta} J^\ast  \boldsymbol{\beta}} = 1}
			\frac{1}{n} \mathbf{d}^\top J^\ast  \boldsymbol{\beta}
			\\& \overset{(i)}{\leq} 
			\max_{\sqrt{\boldsymbol{\beta} J^\ast  \boldsymbol{\beta}} = 1}
			\frac{1}{n} \sqrt{\mathbf{d}^\top J^\ast  \mathbf{d}}
			\sqrt{\boldsymbol{\beta} J^\ast  \boldsymbol{\beta}}
			\\& \overset{(ii)}{=}
			\frac{1}{n} \sqrt{\mathbf{d}^\top J^\ast  \mathbf{d}}    ,
		\end{align*}
		where step $(i)$ uses the Cauchy-Schwarz inequality with the inner product $\langle x, y \rangle_{J^\ast} = x^\top J^\ast y$,
		and step $(ii)$ uses the norm constraint.
		By the Cauchy-Schwarz inequality, the maximum displayed in the last term is obtained by a unit vector (with respect to $\langle \cdot, \cdot \rangle_{J^\ast}$) parallel to $\mathbf{d}$. Therefore we have
		\begin{equation*}
			\begin{bmatrix} \mathbf{a} \\ \mathbf{\theta} \end{bmatrix}
			= \frac{1}{\sqrt{\mathbf{d}^\top  J^\ast \mathbf{d} }}\mathbf{d},
		\end{equation*}
		which only involves solving one linear equation.
	}

@@

			V^\ast = n^{-2}\begin{bmatrix}
				J \widetilde{C}^\top \mathbf{y}\mathbf{y}^\top \widetilde{C} J & J \widetilde{C}^\top \widetilde{\mathbf{y}} \widetilde{\mathbf{y}}^\top \widetilde{Z} \\
				\widetilde{Z}^\top \widetilde{\mathbf{y}} \widetilde{\mathbf{y}}^\top \widetilde{C} J &   \widetilde{Z}^\top \widetilde{\mathbf{y}} \widetilde{\mathbf{y}}^\top \widetilde{Z}

@@

				&\argmax \limits_{\boldsymbol{\xi}^* \in \mathbb{R}^{MK+p}} \; \frac{\boldsymbol{\xi}^{*\top} V^\ast \boldsymbol{\xi}^*}{\boldsymbol{\xi}^{*\top}(J^\ast+\Lambda \ddot{J}^\ast) \boldsymbol{\xi}^*} , \quad \textit{or equivalently}, \\ 
				&\argmax \limits_{\boldsymbol{\xi}^* \in \mathbb{R}^{MK+p}} \;  \boldsymbol{\xi}^{*\top} V^\ast \boldsymbol{\xi}^* \quad \textnormal{subject to}  \quad \boldsymbol{\xi}^{*\top} (J^\ast+\Lambda \ddot{J}^\ast) \boldsymbol{\xi}^*=1.

@@

	To solve \eqref{eq: Regularized generalized rayleigh quotient equation} in practice, we first perform a Choleski factorization: $L L^\top = J^\ast + \Lambda \ddot{J}^\ast$, where $L$ is a lower triangular matrix, and then reduce it to maximizing the classical Rayleigh quotient problem:
	\begin{equation*}
		\max \limits_{\mathbf{e} \in \mathbb{R}^{MK + p}} \, \frac{\mathbf{e}^\top E \mathbf{e}}{\mathbf{e}^\top \mathbf{e}} \; \Longleftrightarrow \;   \max \limits_{\substack{\mathbf{e} \in \mathbb{R}^{MK+p}}} \;  \mathbf{e}^\top E \mathbf{e} \quad \textnormal{subject to} \quad \Vert \mathbf{e} \Vert^2 = 1
	\end{equation*}
	through the transformations $\mathbf{e} = L^\top \boldsymbol{\xi}^*$ and $E = L^{-1} V^\ast L^{\top-1}$. Once $\mathbf{e}$ is obtained as the first eigenvector of $E$, we set $\boldsymbol{\xi}^* = L^{\top-1} \mathbf{e}$ and $\widehat{\boldsymbol{\xi}}[\mathbf{t}] = B^*[\mathbf{t}]^\top \boldsymbol{\xi}^*$.

@@

	\begin{algorithm}[b!]
		\caption{Hybrid PLS (under construction)}
		\label{alg:main}
		\begin{algorithmic}[1]
			\Require 
			Response $\mathbf{y} \in \mathbb{R}^n$,
			%functional predictor evalulation points $\mathbf{t}_1, \ldots, \mathbf{t}_m \in [0,1]^K,$
			hybrid predictors approximation coefficient
			$\widetilde{C}^* \ \in \mathbb{R}^{n \times (MK+p)} $ \eqref{eq: basis approximation of W},
			Gram matrix with respect to basis functions
			$J^\ast \in \mathbb{R}^{(MK+p) \times (MK+p)} $
			\eqref{def:J_and_J_star},
			Gram matrix with respect to second derivatives of the basis functions 
			$\ddot{J}^\ast \in \mathbb{R}^{(MK+p) \times (MK+p)}$
			\eqref{def:J_dotdot_ast},
			regularization matrix $\Lambda  \in \mathbb{R}^{(MK+p) \times (MK+p)}$ \eqref{def:Lambda}
			%%%%%%%%%%%%%%%%%%%%%
			\State
			$\widetilde{C}^{*[1]} \leftarrow \widetilde{C}^*$
			\For{ l = 1, 2, \ldots  $T$}
			\State $V^\ast \leftarrow n^{-2} J^\ast \widetilde{C}^{*\top} \widetilde{\mathbf{y}} \widetilde{\mathbf{y}}^\top \widetilde{C}^{*} J^\ast$
			\Comment{ \eqref{def:V_star}}
			\State 	$\widehat{\boldsymbol{\xi}}_l^\ast \leftarrow \argmax_{\boldsymbol{\xi}^* \in \mathbb{R}^{MK+p}} \;  \boldsymbol{\xi}^{*\top} V^\ast \boldsymbol{\xi}^* \quad \textnormal{subject to}  \quad \boldsymbol{\xi}^{*\top} (J^\ast+\Lambda \ddot{J}^\ast) \boldsymbol{\xi}^*=1
			$
			\Comment{ \eqref{eq: Regularized generalized rayleigh quotient equation}} 
			\State $\widehat{\boldsymbol{\rho}}_l
			\leftarrow 
			\widetilde{C}^* J^\ast  \widehat{\boldsymbol{\xi}}_l^\ast
			$
			\Comment{$\in \mathbb{R}^n$; \eqref{compute_pls_score}}
			\EndFor
			\State Output a
		\end{algorithmic}
	\end{algorithm}

@@

>>>>

	\subsubsection{Finite-basis approximation of the predictors and regression coefficients}\label{sec: finite basis approximation}
In practice, the functional predictors $X^{(1)}, \ldots, X^{(K)}$ are not observed as complete curves, but only through their values at a finite set of time points, which may differ across the $K$ predictors. Thus, each observation is represented by a finite-dimensional vector $X_i^{(k)}(\mathbf{t}_k) \in \mathbb{R}^{|\mathbf{t}_k|}$, where $\mathbf{t}_k \subset \mathcal{T}_k$. The first step in functional data analysis typically involves reconstructing the underlying functional form from these discrete measurements, commonly by assuming that sample paths lie in a finite-dimensional space spanned by basis functions.
	\paragraph{Basis functions.}
For $k = 1, \ldots, K$, let $\Phi^{(k)}:= \{\phi^{(k)}_1, \ldots, \phi^{(k)}_M\}$, with $M \geq 1$, be an $M$-dimensional basis in $L^2(\mathcal{T}_k)$. In our proposed hybrid PLS regression algorithm, this basis system is used to approximate both the predictors and the PLS components. Additionally, it serves as a regularization mechanism, borrowing strength across components and helping to prevent overfitting.

	\paragraph{Basis approximation of the hybrid predictor.}	
	For $k = 1, \ldots, K$  and $i = 1, \ldots, n$, 	
	 we approximate each functional predictor within the span of the basis functions $\Phi^{(k)}$ :
\begin{equation*} 
	X_i^{(k)} \approx  \sum_{m=1}^M   c_{im}^{(k)} \phi_m^{(k)} ,
\end{equation*}
where the coefficients $c_{im}^{(k)}$'s satisfy
	\begin{equation*}
	X_i^{(k)}(t_k) = \sum_{m=1}^M c_{im}^{(k)} \phi^{(k)}_{m}(t_k)~\text{for all}~t_k \in \mathbf{t}_k \subset \mathcal{T}_k,
\end{equation*}

	The number of basis functions $M$ for each functional predictor can be set to a sufficiently large value (e.g., 15 or 20) to incorporate a rich variety of patterns that functional predictors may exhibit. Note that there is no need to carefully choose different $M$ values for different functional predictors, as an appropriate penalization scheme will be introduced in the later section to regularize each functional predictor to achieve a desired smoothness (see Section \ref{sec: Regularization of PLS Components}).

	\paragraph{Basis approximation of the PLS component.}
	We can also approximate the functional part of the PLS component in terms of the same basis functions: $\psi^{(k)}(t_k) = \sum_{m=1}^M a^{(k)}_{m} b^{(k)}_m(t_k) = \mathbf{a}^{(k)\top} \mathbf{b}^{(k)}(t_k)$ and $\psi(\mathbf{t}) = B(\mathbf{t})^\top \mathbf{a} \in \mathbb{R}^K$, where $\mathbf{a}^{(k)} = (a^{(k)}_{1}, \ldots, a^{(k)}_{M})^\top \in \mathbb{R}^M$, and $\mathbf{a}=(\mathbf{a}^{(1)\top}, \ldots, \mathbf{a}^{(K)\top})^\top \in \mathbb{R}^{MK}$, so that $\boldsymbol{\xi}[\mathbf{t}] = (\psi(\mathbf{t})^\top, \boldsymbol{\theta}^\top)^\top = (\mathbf{a}^\top B(\mathbf{t}), \boldsymbol{\theta}^\top)^\top \in \mathbb{R}^{K+p}$. 
	%
	For each $k = 1, \ldots, K$, let $\mathbf{J}^{(k)} \in \mathbb{R}^{M \times M}$ denote the Gram matrix of the basis functions associated with the $k$-th functional predictor, defined by
	\begin{equation}\label{def:gram_basis}
		\mathbf{J}^{(k)} = \left[ \int_{\mathcal{T}k} b_m^{(k)}(t) \, b_n^{(k)}(t) , dt \right]_{m,n=1}^M.
	\end{equation}
	Using this definition, we construct the following block-diagonal matrices:
	\begin{equation}\label{def:J_and_J_star}
		\mathbf{J} = \mathrm{blkdiag}(\mathbf{J}^{(1)}, \cdots, \mathbf{J}^{(K)}) \in \mathbb{R}^{MK \times MK}~\text{and}~\mathbf{J}^\ast=\mathrm{blkdiag}(\mathbf{J}, \mathbf{I}_p) \in \mathbb{R}^{(MK+p) \times (MK+p)}.
	\end{equation}

@@

We now describe the strategy for computing the PLS direction at the $l$-th iteration. For simplicity, we omit the iteration index and denote the residualized response by $Y$, with finite-sample observations given by $\mathbf{y} = (y_1, \ldots, y_n) \in \mathbb{R}^n$. Let $\widetilde{W} \in \mathcal{H}$ denote the residualized, basis-approximated hybrid predictor, and let $\widetilde{\mathbf{C}} \in \mathbb{R}^{n \times (MK)}$ denote the basis coefficients of its functional part in the finite-sample observations. Our strategy seeks the direction $\boldsymbol{\xi} \in \mathcal{H}$, lying in the span of the basis $B^*[\mathbf{t}]$, that maximizes the empirical covariance between the resulting PLS component $\langle \widetilde{W}, \boldsymbol{\xi} \rangle_{\mathcal{H}}$ and the response.

For practical computation, 
let us first denote
\begin{equation*}
\mathbf{b} := \mathbf{y}^\top
[ \widetilde{\mathbf{C}} \, \mathbf{J} \, \mathbf{Z} ].
\end{equation*}
Then we obtain $\mathbf{d}$ by solving a linear system
\begin{equation*}
	\mathbf{J}^\ast \mathbf{d} =
	\mathbf{b}^\top.
\end{equation*}  
Then we obtain the coefficient for the PLS direction by normalize it with respect to the basis $B^*[\mathbf{t}]$:
\begin{equation}\label{compute:pls_direction_coeff}
	\begin{bmatrix} \mathbf{a} \\ \mathbf{\theta} \end{bmatrix}
	= \frac{1}{\sqrt{\mathbf{d}^\top \mathbf{J}^\ast \mathbf{d}}} \mathbf{d}.
\end{equation}
The $l$-th  PLS score for the 
$i$th observation is then computed as the inner product  with respect to the basis $B^*[\mathbf{t}]$:
\begin{equation*}
	\widehat{\rho}_{i}
	= 
	[ \widetilde{\mathbf{C}}  \, \mathbf{Z} ] \, \mathbf{J}^\ast \begin{bmatrix} \mathbf{a} \\ \mathbf{\theta} \end{bmatrix}
\end{equation*}
$ \equiv \widehat{\rho}_{il}$ can be obtained as $\widehat{\rho}_i = \langle \mathbf{W}_i, \widehat{\boldsymbol{\xi}} \rangle_\mathcal{H}$.
we first construct the following matrix:
	\begin{equation}\label{def:V_star}
	\mathbf{V}^\ast 
	:= 
	n^{-2}\begin{bmatrix}
\mathbf{J}
\widetilde{ \mathbf{C} }^\top \mathbf{y}\mathbf{y}^\top \widetilde{ \mathbf{C} } \mathbf{J} 
& 
\mathbf{J} \widetilde{ \mathbf{C} }^\top 
\mathbf{y}
\mathbf{y}^\top 
\mathbf{Z} 
\\
\mathbf{Z}^\top 
\mathbf{y}
\mathbf{y}^\top \widetilde{ \mathbf{C} }
\mathbf{J} 
&   
\mathbf{Z}^\top 
\mathbf{y}
\mathbf{y}^\top 
\mathbf{Z}
	\end{bmatrix} = n^{-2} \mathbf{J}^\ast \widetilde{ \mathbf{C} }^{*\top} \widetilde{\mathbf{y}} \widetilde{\mathbf{y}}^\top \widetilde{ \mathbf{C} }^{*} \mathbf{J}^\ast
	\in \mathbb{R}^{(MK+p) \times (MK+p)}.
\end{equation}

%	consider the basis function expansions of the functional predictors $X_i^{(k)}(t_k) = \sum_{m=1}^M c_{im}^{(k)} b^{(k)}_{m}(t_k)$, $i=1, \ldots, n$, and those of the functional parts of the PLS component $\psi^{(k)}(t_k) = \sum_{m=1}^M a^{(k)}_{m} b^{(k)}_m(t_k)$, $k=1, \ldots, K$. Let 
	\begin{proposition} \label{Proposition: generalized rayleigh quotient}
		Let $\operatorname{span}(B^\ast[\mathbf{t}])$ 
		be a linear subspace of $\mathcal{H}$ spanned by the basis functions.
		  For any unit-length direction  $\boldsymbol{\xi} \in \operatorname{span}(B^\ast[\mathbf{t}])$,
expressed as $
\boldsymbol{\xi} 
= 
\boldsymbol{\alpha}
B^\ast [\mathbf{t}] $ with $\boldsymbol{\alpha} \in \mathbb{R}^{MK + p}$,
		Let us define
		the empirical and basis-approximated sample covariance as
\begin{equation*}
	\widehat{\textnormal{Cov}} (\langle \widetilde{\mathbf{W}}, \boldsymbol{\xi} \rangle_{\mathcal{H}}, Y )
	:=
	\frac{1}{n} \sum \limits_{i=1}^n \langle \widetilde{W}_i, \boldsymbol{\xi} \rangle_\mathcal{H} y_i
	=
		\frac{1}{n} \sum \limits_{i=1}^n 
	(	\tilde{\mathbf{C}}^\ast_i \mathbf{J}^\ast \boldsymbol{\alpha} ) y_i.
\end{equation*}
Then the coefficients computed in \eqref{compute:pls_direction_coeff}  $\widehat{\textnormal{Cov}} (\langle \widetilde{\mathbf{W}}, \boldsymbol{\xi} \rangle_{\mathcal{H}}, Y )$
over all unit vectors in $\operatorname{span}(B^\ast[\mathbf{t}])$.
In a more formal way:
	\begin{equation}
	\begin{bmatrix} \mathbf{a} \\ \mathbf{\theta} \end{bmatrix}
	B^*[\mathbf{t}] \in \arg \max_{
 \|\boldsymbol{\xi}  \| = 1, 
	 \boldsymbol{\xi} \in \mathrm{span}(B^*[\mathbf{t}])
	}
		\widehat{\textnormal{Cov}} (\langle \widetilde{\mathbf{W}}, \boldsymbol{\xi} \rangle_{\mathcal{H}}, Y ).
\end{equation}

	The proof of Proposition \ref{Proposition: generalized rayleigh quotient} is given in the Appendix \ref{section:proof:eigenproblem_derivation}.

@@

			\mathbf{V}^\ast = n^{-2}\begin{bmatrix}
				J \widetilde{C}^\top \mathbf{y}\mathbf{y}^\top \widetilde{C} J & J \widetilde{C}^\top \widetilde{\mathbf{y}} \widetilde{\mathbf{y}}^\top \mathbf{Z} \\
				\mathbf{Z}^\top \widetilde{\mathbf{y}} \widetilde{\mathbf{y}}^\top \widetilde{C} J &   \mathbf{Z}^\top \widetilde{\mathbf{y}} \widetilde{\mathbf{y}}^\top \mathbf{Z}

@@

				&\argmax \limits_{\boldsymbol{\xi}^* \in \mathbb{R}^{MK+p}} \; \frac{\boldsymbol{\xi}^{*\top} \mathbf{V}^\ast \boldsymbol{\xi}^*}{\boldsymbol{\xi}^{*\top}(J^\ast+\Lambda \ddot{J}^\ast) \boldsymbol{\xi}^*} , \quad \textit{or equivalently}, \\ 
				&\argmax \limits_{\boldsymbol{\xi}^* \in \mathbb{R}^{MK+p}} \;  \boldsymbol{\xi}^{*\top} \mathbf{V}^\ast \boldsymbol{\xi}^* \quad \textnormal{subject to}  \quad \boldsymbol{\xi}^{*\top} (J^\ast+\Lambda \ddot{J}^\ast) \boldsymbol{\xi}^*=1.

@@

	To solve \eqref{eq: Regularized generalized rayleigh quotient equation} in practice, we first perform a Choleski factorization: $L L^\top = J^\ast + \Lambda \ddot{J}^\ast$, where $L$ is a lower triangular matrix, and then reduce it to maximizing the classical Rayleigh quotient problem:
	\begin{equation*}
		\max \limits_{\mathbf{e} \in \mathbb{R}^{MK + p}} \, \frac{\mathbf{e}^\top E \mathbf{e}}{\mathbf{e}^\top \mathbf{e}} \; \Longleftrightarrow \;   \max \limits_{\substack{\mathbf{e} \in \mathbb{R}^{MK+p}}} \;  \mathbf{e}^\top E \mathbf{e} \quad \textnormal{subject to} \quad \Vert \mathbf{e} \Vert^2 = 1
	\end{equation*}
	through the transformations $\mathbf{e} = L^\top \boldsymbol{\xi}^*$ and $E = L^{-1} \mathbf{V}^\ast L^{\top-1}$. Once $\mathbf{e}$ is obtained as the first eigenvector of $E$, we set $\boldsymbol{\xi}^* = L^{\top-1} \mathbf{e}$ and $\widehat{\boldsymbol{\xi}}[\mathbf{t}] = B^*[\mathbf{t}]^\top \boldsymbol{\xi}^*$.

@@

	\begin{algorithm}[b!]
		\caption{Hybrid PLS (under construction)}
		\label{alg:main}
		\begin{algorithmic}[1]
			\Require 
			Response $\mathbf{y} \in \mathbb{R}^n$,
			%functional predictor evalulation points $\mathbf{t}_1, \ldots, \mathbf{t}_m \in [0,1]^K,$
			hybrid predictors approximation coefficient
			$\widetilde{C}^* \ \in \mathbb{R}^{n \times (MK+p)} $ \eqref{eq: basis approximation of W},
			Gram matrix with respect to basis functions
			$J^\ast \in \mathbb{R}^{(MK+p) \times (MK+p)} $
			\eqref{def:J_and_J_star},
			Gram matrix with respect to second derivatives of the basis functions 
			$\ddot{J}^\ast \in \mathbb{R}^{(MK+p) \times (MK+p)}$
			\eqref{def:J_dotdot_ast},
			regularization matrix $\Lambda  \in \mathbb{R}^{(MK+p) \times (MK+p)}$ \eqref{def:Lambda}
			%%%%%%%%%%%%%%%%%%%%%
			\State
			$\widetilde{C}^{*[1]} \leftarrow \widetilde{C}^*$
			\For{ l = 1, 2, \ldots  $T$}
			\State $\mathbf{V}^\ast \leftarrow n^{-2} J^\ast \widetilde{C}^{*\top} \widetilde{\mathbf{y}} \widetilde{\mathbf{y}}^\top \widetilde{C}^{*} J^\ast$
			\Comment{ \eqref{def:V_star}}
			\State 	$\widehat{\boldsymbol{\xi}}_l^\ast \leftarrow \argmax_{\boldsymbol{\xi}^* \in \mathbb{R}^{MK+p}} \;  \boldsymbol{\xi}^{*\top} \mathbf{V}^\ast \boldsymbol{\xi}^* \quad \textnormal{subject to}  \quad \boldsymbol{\xi}^{*\top} (J^\ast+\Lambda \ddot{J}^\ast) \boldsymbol{\xi}^*=1
			$
			\Comment{ \eqref{eq: Regularized generalized rayleigh quotient equation}} 
			\State $\widehat{\boldsymbol{\rho}}_l
			\leftarrow 
			\widetilde{C}^* J^\ast  \widehat{\boldsymbol{\xi}}_l^\ast
			$
			\Comment{$\in \mathbb{R}^n$; \eqref{compute_pls_score}}
			\EndFor
			\State Output a
		\end{algorithmic}
	\end{algorithm}

@@

	\section{old notations}
